Lab 9: Multinomial and Ordered Probit Models: Answers

### 2023-03-23 Lab 9  Multinom_orderedprob/Lab9_Multinom_OrderedProbit.Rmd
Add files via upload

Generate a cumulative predicted probability plot varying the prestige of the respondent's occupation.

@@

#Generate a cumulative predicted probability plot.
table(hw$male); table(hw$white); mean(hw$ed)

hw_2 <- expand.grid(_____,
                      prst = seq(from = min(hw$prst), to = max(hw$prst), length.out = 2293))

@@

plot_data <- data.frame(hw$prst, cumul_probs)

#Generate a cumulative predicted probability plot.
ggplot(data=plot_data, aes(x=hw.prst))+
      ______

>>>>

Generate a cumulative predicted probability plot varying the age of the respondent.

@@

#Compute age quartiles and create hypothetical age values between min and max
age_quarts <- with(hw,c(min(age,na.rm = TRUE), max(age,na.rm=TRUE)))

age_rug <- with(hw,seq(age_quarts[1], age_quarts[2], length.out=100))

#Finding modal values
table(hw$male); table(hw$white); mean(hw$ed); mean(hw$prst)
hw_2 <- expand.grid(male = 1, #modal value
                      white = 1, #modal value
                      ed=mean(hw$ed),
                      prst=mean(hw$prst),
                      age = age_rug)

@@

plot_data <- data.frame(age_rug, cumul_probs)

#Generate a cumulative predicted probability plot varying by age
ggplot(data=plot_data, aes(x=age_rug))+
      ______

### 2023-03-23 Lab 9  Multinom_orderedprob/Lab9_Multinom_OrderedProbit_Answers.Rmd
Add files via upload

#Generate a cumulative predicted probability plot.
table(hw$male); table(hw$white); mean(hw$ed)
hw_2 <- expand.grid(male = 1, #modal value
                      white = 1, #modal value
                      ed=mean(hw$ed),
                      age=mean(hw$age),
                      prst = seq(from = min(hw$prst), to = max(hw$prst), length.out = 2293))

probs <- predict(warm_model, hw_2, type = "prob") #type = is different than for binomial model
head(rowSums(probs$fit)) #just to double check
cumul_probs <- t(apply(probs$fit, 1, cumsum)) #get the cumulative sum of the rows
#transpose into columns for plotting and merging with values of gpa
head(cumul_probs)
plot_data <- data.frame(hw$prst, cumul_probs)

#plot data
ggplot(data=plot_data, aes(x=hw.prst))+
    geom_smooth(data=plot_data, aes(x=hw.prst, y=Strongly.disagree), se=FALSE)+
    geom_smooth(data=plot_data, aes(x=hw.prst, y=Disagree), se=FALSE)+
    geom_smooth(data=plot_data, aes(x=hw.prst, y=Agree), se=FALSE)+
      geom_smooth(data=plot_data, aes(x=hw.prst, y=Strongly.agree), se=FALSE)+
    geom_ribbon(data=plot_data, aes(ymin=0, ymax=Strongly.disagree, fill="Strongly.disagree"))+
    geom_ribbon(data=plot_data, aes(ymin=Strongly.disagree, ymax=Disagree,
                                    fill="Disagree"))+
    geom_ribbon(data=plot_data, aes(ymin=Disagree, ymax=Agree,
                                    fill="Agree"))+
         geom_ribbon(data=plot_data, aes(ymin=Agree, ymax=Strongly.agree,
                                    fill="Strongly.agree"))+
    labs(x="Values of prestige of occupation", y="Cumulative Predicted Probability")+
    scale_fill_discrete(name="Hard Work")+
    theme_bw()

>>>>

#Generate a cumulative predicted probability plot.

#Compute age quartiles and create hypothetical age values between min and max
age_quarts <- with(hw,c(min(age,na.rm = TRUE), max(age,na.rm=TRUE)))

age_rug <- with(hw,seq(age_quarts[1], age_quarts[2], length.out=100))

#Finding modal values
table(hw$male); table(hw$white); mean(hw$ed); mean(hw$prst)
hw_2 <- expand.grid(male = 1, #modal value
                      white = 1, #modal value
                      ed=mean(hw$ed),
                      prst=mean(hw$prst),
                      age = age_rug)

probs <- predict(warm_model, hw_2, type = "prob") #type = is different than for binomial model
head(rowSums(probs$fit)) #just to double check
cumul_probs <- t(apply(probs$fit, 1, cumsum)) #get the cumulative sum of the rows
#transpose into columns for plotting and merging with values of gpa
head(cumul_probs)
plot_data <- data.frame(age_rug, cumul_probs)

#plot data
ggplot(data=plot_data, aes(x=age_rug))+
    geom_smooth(data=plot_data, aes(x=age_rug, y=Strongly.disagree), se=FALSE)+
    geom_smooth(data=plot_data, aes(x=age_rug, y=Disagree), se=FALSE)+
    geom_smooth(data=plot_data, aes(x=age_rug, y=Agree), se=FALSE)+
      geom_smooth(data=plot_data, aes(x=age_rug, y=Strongly.agree), se=FALSE)+
    geom_ribbon(data=plot_data, aes(ymin=0, ymax=Strongly.disagree, fill="Strongly.disagree"))+
    geom_ribbon(data=plot_data, aes(ymin=Strongly.disagree, ymax=Disagree,
                                    fill="Disagree"))+
    geom_ribbon(data=plot_data, aes(ymin=Disagree, ymax=Agree,
                                    fill="Agree"))+
         geom_ribbon(data=plot_data, aes(ymin=Agree, ymax=Strongly.agree,
                                    fill="Strongly.agree"))+
    labs(x="Age of respondent", y="Cumulative Predicted Probability")+
    scale_fill_discrete(name="Warm Relationship w/ Child")+
    theme_bw()
